---
title: "Untitled"
output: html_document
date: "2023-06-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## by circlize
```{r}
library(circlize)
library(tidyverse)

#dataForPlot1=mtcars %>% rownames_to_column()
dataForPlot1=data.frame(mtcars)

dataForPlot1$gear=as.factor(dataForPlot1$gear)
```

### First make sure all features are correctly and consistently ordered
### Order by value (for example, wt)
```{r}
dataForPlot1=dataForPlot1 %>% group_by(gear) %>% mutate(RankedPosition=rank(wt,ties.method="first"))
#save back rownames
dataForPlot1=data.frame(dataForPlot1)
row.names(dataForPlot1)=row.names(mtcars)

#Now all data need to be ordered by this RankedPosition

```

### Order by heatmap (for example, heatmap of mpg, cyl, disp)

CELL_META$subset: the index of subset used in current sector (split)
CELL_META$row_order: the row ordering of the sub-matrix in the current sector after clustering. 

See the code below. Use **y = row_mean[CELL_META$subset]** to subset row_mean data in current sector, and then use **y = y[CELL_META$row_order]** to reorder the subsetted vector.

See this link for detail: https://jokergoo.github.io/circlize_book/book/circos-heatmap.html#with-other-tracks

```{r}
col_fun1 = colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))

circos.clear()
circos.heatmap(scale(dataForPlot1[,c("mpg","cyl","disp")]), split = dataForPlot1$gear,col=col_fun1,rownames.side = "out")
row_mean = rowMeans(scale(dataForPlot1[,c("mpg","cyl","disp")]))
circos.track(ylim = range(row_mean), panel.fun = function(x, y) {
    y = row_mean[CELL_META$subset]
    y = y[CELL_META$row_order]
    circos.lines(CELL_META$cell.xlim, c(0, 0), lty = 2, col = "grey")
    circos.points(seq_along(y) - 0.5, y, col = ifelse(y > 0, "red", "blue"))
}, cell.padding = c(0.02, 0, 0.02, 0))

```

### If you need add something before heatmap (for example, a barplot or a color bar)

Similarly, if the points track is put as the first track, the layout should be initialized in advance.

```{r}
circos.heatmap.initialize(mat1, split = split)
# This is the same as the previous example
circos.track(ylim = range(row_mean), panel.fun = function(x, y) {
    y = row_mean[CELL_META$subset]
    y = y[CELL_META$row_order]
    circos.lines(CELL_META$cell.xlim, c(0, 0), lty = 2, col = "grey")
    circos.points(seq_along(y) - 0.5, y, col = ifelse(y > 0, "red", "blue"))
}, cell.padding = c(0.02, 0, 0.02, 0))
circos.heatmap(mat1, col = col_fun1) # no need to specify 'split' here


```

### Make category by colors and also add names outside
```{r}


bucketColors=pals::kelly(length(unique(dataForPlot1$gear))+2)[-c(1:2)]
col_anno_sector = structure(bucketColors, names = as.character(unique(dataForPlot1$gear)))

circos.clear()
circos.heatmap(dataForPlot1[,"gear",drop=FALSE], 
               col = col_anno_sector,  split=dataForPlot1$gear,
               rownames.cex=0.7,cluster=FALSE,rownames.side="outside",
               track.height = 0.05,track.margin=c(0, 0))
```

### Add barplot by colors
```{r}
circos.track(selectedProteinsMergedTable$bucket, 
             x = selectedProteinsMergedTable$BarPos, 
             y = selectedProteinsMergedTable$beta, 
             track.height = 0.2,track.margin=c(0.0, 0.0),
             bg.border="white",
    panel.fun = function(x, y,colour) {
    circos.barplot(value=y, pos=x-0.5, col =  ifelse(y > 0, barplotColors[1], barplotColors[2]),border="white")
})
```

### Add name for each track
```{r}
circos.track(track.index = get.current.track.index(), panel.fun = function(x, y) {
     if(CELL_META$sector.numeric.index == 4) { # the last sector
        cn = "Max correlation"
        n = 1
        circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
            1:n - 0.5, cn, 
            cex = 0.5, adj = c(0, 0.5), facing = "inside")
     }
}, bg.border = NA)

#for heatmap
circos.track(track.index = get.current.track.index(), panel.fun = function(x, y) {
     if(CELL_META$sector.numeric.index == 4) { # the last sector
        cn = rev(heatmapNameExpression) #please note the order needs to be reversed as heatmap is from top to bottom and 1:n is from bottom to top
        n = length(cn)
        circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"),
            1:n - 0.5, cn,
            cex = 0.5, adj = c(0, 0.5), facing = "inside")
     }
}, bg.border = NA)
```

### Add links or lines
See this link: https://jokergoo.github.io/circlize_book/book/circos-heatmap.html
```{r}
for(i in seq_len(nrow(df_link))) {
    circos.heatmap.link(df_link$from_index[i],
                        df_link$to_index[i],
                        col = rand_color(1))
}
```



```{r}


bucketColors=kelly(length(unique(selectedProteinsMergedTable$bucket))+2)[-c(1:2)]
col_anno_sector = structure(bucketColors, names = unique(selectedProteinsMergedTable$bucket))
col_anno_sector

barplotColors=c("red","green")

selectedProteinsMergedTable=selectedProteinsMergedTable %>% group_by(bucket) %>% mutate(BarPos=rank(beta,ties.method="first"))

dataForPlot1=data.frame(bucket=selectedProteinsMergedTable$bucket)
row.names(dataForPlot1)=selectedProteinsMergedTable$Protein

circos.clear()
circos.heatmap(dataForPlot1, 
               col = col_anno_sector,  split=selectedProteinsMergedTable$bucket,
               rownames.cex=0.7,cluster=FALSE,rownames.side="outside",
               track.height = 0.05,track.margin=c(0, 0))
circos.track(selectedProteinsMergedTable$bucket, 
             x = selectedProteinsMergedTable$BarPos, 
             y = selectedProteinsMergedTable$beta, 
             track.height = 0.2,track.margin=c(0.0, 0.0),
             bg.border="white",
    panel.fun = function(x, y,colour) {
    circos.barplot(value=y, pos=x-0.5, col =  ifelse(y > 0, barplotColors[1], barplotColors[2]),border="white")
})

```

## Other important parameters
### Before making figure
```{r}
#leave space for names
circos.par(gap.after = c(2, 2, 2, 24))
```

### When making figure
```{r}
               col=featureTypeColor;
               col=col_fun_feature;
               rownames.side = "out";
               track.height = 0.02;
               track.margin=c(0,0)
```



