---
title: "Colors in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Colors in ggplot2

[http://www.cookbook-r.com/Graphs/Colors\_(ggplot2)/](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/){.uri}
http://www.sthda.com/english/wiki/ggplot2-colors-how-to-change-colors-automatically-and-manually


```{r}
library(tidyverse)

# Two variables
df <- read.table(header=TRUE, text='
 cond yval
    A 2
    B 2.5
    C 1.6
')

# Three variables
df2 <- read.table(header=TRUE, text='
 cond1 cond2 yval
    A      I 2
    A      J 2.5
    A      K 1.6
    B      I 2.2
    B      J 2.4
    B      K 1.2
    C      I 1.7
    C      J 2.3
    C      K 1.9
')
```

### Concepts

#### Colors: RGB, CMYK, transprency, luminance and saturation




#### fill vs colour

```{r}
ggplot(df, aes(x=cond, y=yval)) + geom_bar(stat="identity", fill="#FF9999", colour="black")

```

#### Mapping variable values to colors: aes or aes_string

```{r}
# Bars: x and fill both depend on cond2
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat="identity")

# fill depends on cond2
ggplot(df2, aes(x=cond1, y=yval)) + 
    geom_bar(aes(fill=cond2,colour=cond1),   # fill depends on cond2
             stat="identity",
             position=position_dodge()) # Put bars side-by-side instead of stacked


```

### Default colors

By default, the colors for discrete scales are evenly spaced around a HSL color circle. For example, if there are two colors, then they will be selected from opposite points on the circle; if there are three colors, they will be 120° apart on the color circle; and so on. The colors used for different numbers of levels are shown here:

![Colors with different levels (N)](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/figure/unnamed-chunk-7-1.png)

#### Color selection and color panels

The default color selection uses scale_fill_hue() and scale_colour_hue().

```{r}
ggplot(df, aes(x=cond, y=yval, colour=cond)) + geom_point(size=2)
ggplot(df, aes(x=cond, y=yval, colour=cond)) + geom_point(size=2) + scale_colour_hue()

```

#### Setting luminance and saturation (chromaticity)

```{r}
# Use luminance=45, instead of default 65
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat="identity") +
    scale_fill_hue(l=40)

# Reduce saturation (chromaticity) from 100 to 50, and increase luminance
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat="identity") +
    scale_fill_hue(c=45, l=80)
```

![Colors with different levels (N), luminance=45](http://www.cookbook-r.com/Graphs/Colors\_(ggplot2)/figure/unnamed-chunk-10-1.png)

#### Change colors manually





### Color functions

```{r}


```

## Other color related packages
https://jmw86069.github.io/colorjam/index.html

```{r}
library(colorjam)
library(pals)

mixOmics::color.jet
```

