---
title: "DESeq2 test and design"
author: "Shilin Zhao"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DESeq2 Design

https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
?results

Different methods of using contrast
* condition effect for genotype I (main effect): 
results(dds, contrast=c("condition","B","A")) #Regular contrast, compare B vs A in condition;

* condition effect for genotype III (the main effect *plus* the interaction term): 
results(dds, contrast=list( c("condition_B_vs_A","genotypeIII.conditionB") )) #list(c(A,B)) to get effect of comprasionA + comprasionB;

* condition effect in genotype III vs genotype II (condition effect is different in III compared to II): results(dds, contrast=list("genotypeIII.conditionB", "genotypeII.conditionB")) #list(A,B) to get effect of comprasionA - comprasionB;

* condition effect for genotypeIII (genotypeIII was NOT in resultsNames(dds)): results(dds, contrast=c("group", "IIIB", "IIIA"))




```{r}
library(DESeq2)
example(results)

#some test
dds <- makeExampleDESeqDataSet(n=100,m=12)
dds$genotype <- factor(rep(rep(c("I","II"),each=3),2))

design(dds) <- ~ genotype + condition + genotype:condition
dds <- DESeq(dds) 
resultsNames(dds)

# Note: design with interactions terms by default have betaPrior=FALSE

# the condition effect for genotype I (the main effect)
results(dds, contrast=c("condition","B","A"))

# the condition effect for genotype II
# this is, by definition, the main effect *plus* the interaction term
# (the extra condition effect in genotype II compared to genotype I).
results(dds, list( c("condition_B_vs_A","genotypeII.conditionB") ))

# the interaction term, answering: is the condition effect *different* across genotypes?
results(dds, name="genotypeII.conditionB")
 





```

