---
title: "Statistic Methods Example"
author: "Shilin Zhao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 3000)
```

## Functions and packages

```{r}
library(knitr)
library(ggplot2)
library(RColorBrewer)
library(reshape2)
library(formattable)
library(rms)
library(htmlTable)

library(cqsR)
source('D:/source/cqsR/R/rmsEasierModel.R')
#devtools::source_url("https://raw.github.com/slzhao/cqsR/master/R/rmsEasierModel.R")

```


## Make Data

```{r}
rawData<-matrix(nrow=200,ncol=10,rnorm(200*10))
row.names(rawData)<-paste0("Sample",1:200)
colnames(rawData)<-paste0("Feature",1:10)
rawData<-data.frame(rawData,FeatureYN1=sample(c(0,1),200,replace = TRUE),FeatureCategory1=sample(c("F","E"),200,replace = TRUE),FeatureCategory2=sample(c("A","B","C","D"),200,replace = TRUE),stringsAsFactors = TRUE)

#make data with missing values
rawDataWithMissing<-rawData

set.seed(123)
temp1<-sample(1:ncol(rawDataWithMissing),200,replace=TRUE)
temp2<-sample(1:nrow(rawDataWithMissing),200,replace=TRUE)
for (i in 1:length(temp1)) {
  rawDataWithMissing[temp2[i],temp1[i]]<-NA
}


```

## Notes

```{r}
#Making formula. Using () to print at the same time
#(formulaForModel<-as.formula(paste0(outVar,"~",paste(varForTable1, collapse=" + "))))
```


##Logistic regression: Model, Table, and Figure

### Model and Table

```{r,comment="",results='asis'}
outVar="FeatureYN1"
varForTable=c("Feature1", "Feature2","Feature3","FeatureCategory1")

nonLinearTest(rawData,outVar,varForTable,modelType ="lrm",returnKable=TRUE)

modelTable(rawData,outVars="FeatureYN1",interestedVars=list(c("Feature1"),c("Feature2","FeatureCategory1")),adjVars="Feature3")
modelTable(rawData,outVars="FeatureYN1",interestedVars=c("Feature1","Feature2","FeatureCategory1"),adjVars="Feature3")

```

### Figure

```{r,comment="",cache=TRUE}
outVar="FeatureYN1"
varForTable=c("Feature1", "Feature2","Feature3","FeatureCategory1")

nonLinearTest(rawData,outVar,varForTable,modelType ="lrm")

#None of them has non-linear term. But will add nonlinear in the model as an example
#(formulaForModel<-as.formula(paste0(outVar,"~",paste(varForTable, collapse=" + "))))
formulaForModel<-as.formula(FeatureYN1~rcs(Feature1,3)+Feature2+Feature3+FeatureCategory1)
dataForModel<-rawData[,c(outVar,varForTable)]
ddist <- datadist(dataForModel)
options(datadist='ddist')

modelResult <- lrm(formulaForModel, data=rawData)
print(modelResult)
#showOddsRatio2(modelResult)
plot(Predict(modelResult))


#extract p value from results
#http://r.789695.n4.nabble.com/Extracting-P-values-from-the-lrm-function-in-the-rms-library-td2260999.html
pnorm(abs(modelResult$coef/sqrt(diag(modelResult$var))),lower.tail=F)*2 

#more plot
plot(Predict(modelResult,fun=plogis))

plot(Predict(modelResult,Feature1,fun=plogis))
plot(Predict(modelResult,Feature1=sort(dataForModel$Feature1),fun=plogis))


temp=Predict(modelResult,Feature1=sort(dataForModel$Feature1),fun=plogis)
str(temp)
p=rms:::ggplot.Predict(temp,ylim=c(0,1))
#same thing
#p=ggplot(temp,aes(x=Feature1,y=yhat),ylim=c(0,1))+geom_line()+geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.05)

p+geom_point(aes(x=Feature1,y=FeatureYN1),data=dataForModel)

```


##Logistic regression: ROC curve and best cutoff

### Simple one variable model

https://stackoverflow.com/questions/16347507/obtaining-threshold-values-from-a-roc-curve

Or using OptimalCutpoints package which provides many algorithms to find an optimal thresholds.
Or coords function in pROC package

```{r,results='asis',cache=TRUE}
source("D:\\source\\r_cqs\\myPkg\\R\\logisticRegressionAndRocFunctions.R")

dataForModel=rawData
temp=makeModelAndTable(dataForModel,labelVar="FeatureYN1",xVar="Feature1")
kable(temp)
```

### Complicated Model

```{r,cache=TRUE}
source("D:\\source\\r_cqs\\myPkg\\R\\logisticRegressionAndRocFunctions.R")

otherVar=c("FeatureYN1")
xVarList<-list(
		"Feature1",
		c("Feature2","FeatureCategory1")
)
otherVar=c("Feature3") #like age and gender
plotModelRocAdjusted(dataForModel,
		xVarList=xVarList,outVar=outVar,otherVar=otherVar,
		verbose=TRUE)


```




## Linear regression

```{r,comment=""}
outVar="Feature1"
varForTable=c("FeatureYN1", "Feature2","Feature3","FeatureCategory1")

nonLinearTest(rawData,outVar,varForTable,modelType ="ols")



modelTable(rawData,outVars="Feature1",interestedVars=list(c("FeatureYN1"),c("Feature2","FeatureCategory1")),adjVars="Feature3",modelType ="ols")
modelTable(rawData,outVars="Feature1",interestedVars=c("FeatureYN1","Feature2","FeatureCategory1"),adjVars="Feature3",modelType ="ols")



```

## Survival Model

### KM Curves

```{r}
source("D:\\source\\r_cqs\\myPkg\\R\\survivalCurveTable.R")

set.seed(123)
rawData$FeatureAbs1<-abs(rnorm(nrow(rawData))*100)

fit1 <- survfit(Surv(FeatureAbs1, FeatureYN1) ~ FeatureCategory1,data=rawData)
kmplot(fit1,col.surv=1:2,lwd.surv=3,grid=FALSE,las=1,pValue=TRUE,cex.axis=1.2,loc.legend="topright")
print(fit1)

```

### Model
```{r}
ddist <- datadist(rawData)
options(datadist='ddist')

varForSurvivalModel1=c("FeatureCategory1","Feature2")
(fmla<-as.formula(paste0("Surv(FeatureAbs1, FeatureYN1)~",paste0(varForSurvivalModel1,collapse="+"))))
fmi=cph(fmla,data=rawData)

print(fmi)
#showOddsRatio2(fmi)
plot(Predict(fmi))

```



##Missing values

### Table 0: Missing Data Describe
```{r,results='asis',cache=TRUE}
varForTable=colnames(rawDataWithMissing)
s=describe(rawDataWithMissing[,varForTable])
html(s, exclude1=FALSE,  what=c('%'),digits=3, prmsd=TRUE)
```




### Missing Value Imputation
```{r,cache=TRUE}
dataImputated <- aregImpute(~Feature1+Feature2+Feature3+
				Feature4+Feature5+Feature6+Feature7+Feature8+Feature9+Feature10+
				FeatureYN1+FeatureCategory1+FeatureCategory2,
		n.impute=10, data=rawDataWithMissing, nk=0,x=TRUE)
print(dataImputated)


#Values after imputation
names(dataImputated$imputed)
tail(dataImputated$imputed$Feature1)

```



### Model after Missing Value Imputation
```{r,cache=TRUE}
fmi <- fit.mult.impute(FeatureYN1~Feature1+Feature2+Feature3+FeatureCategory2,
		lrm, dataImputated,  
#		subset=which(rawDataWithMissing$FeatureCategory1=="E"),
		data=rawDataWithMissing)
print(fmi)

```


##Some notes for Hmisc/rms package
```{r,eval=FALSE}
library(scales)
ggplot(Predict(gls1))
print("Exp")

p<-ggplot(Predict(gls1),flipxdiscrete=FALSE,sepdiscrete="list",ylim(3,6))
p[[1]]
p[[1]]+scale_y_continuous(trans="exp",limits = c(3,6))

exp_format <- function() {
     function(x) format(exp(x),digits = 2) 
}
p[[1]]+scale_y_continuous(trans="exp",limits = c(3,6),labels = exp_format())
p[[1]]+scale_y_continuous(labels = exp_format())


p[[2]]
p[[2]]+scale_x_continuous(trans="exp")


ggplot(Predict(gls1),flipxdiscrete=FALSE,addlayer=scale_y_continuous(trans="exp",labels=function(x) round(exp(x),2)))
ggplot(Predict(gls1),flipxdiscrete=FALSE)+coord_trans(y = "exp")

```


## Two-Way ANOVA with Repeated Measures
https://biostats.w.uib.no/factorial-repeated-measures-anova-two-way-repeated-measures-anova/
http://psych.wisc.edu/moore/Rpdf/610-R9_Within2way.pdf
https://datascienceplus.com/two-way-anova-with-repeated-measures/

```{r,cache=TRUE}
rat.weight <- c(166,168,155,159,151,166,170,160,162,153,220,230,223,233,229,262,274,267,283,274,261,275,264,280,282,343,354,351,359,349,297,311,305,315,303,375,399,388,405,395)
rat.strain <- as.factor(rep(c(rep("strainA",5),rep("strainB",5)),4))
rat.ID <- as.factor(rep(c("rat01","rat02","rat03","rat04","rat05","rat06","rat07","rat08","rat09","rat10"),4))
time.point <- as.factor(c(rep("week08",10), rep("week12",10), rep("week16",10), rep("week20",10)))
my.dataframe <- data.frame(rat.ID,rat.strain,time.point,rat.weight)
head(my.dataframe)

with(my.dataframe, interaction.plot(time.point, rat.ID, rat.weight, pch=c(rep(1,5), rep(2,5)), type="b", col=c("black", "red", "blue", "green", "purple", "orange", "grey", "darkblue", "pink", "darkred"), lty= c(1,2,3,4,5,6,7,8,9,10)))

#we need to check for normality of distribution with the Shapiro-Wilk test:
shapiro.test(rat.weight[time.point=="week08" & rat.strain=="strainA"])
```

### Two-Way ANOVA with Repeated Measures in dataset1

```{r,cache=TRUE}
results <- aov(rat.weight~time.point*rat.strain + Error(rat.ID/time.point), data=my.dataframe)
summary(results)
```

### A linear mixed-effects model with nested random effects. As a validation.
The F value is the same as Two-Way ANOVA above

```{r}
library(nlme)

results.lme <- lme(rat.weight~time.point*rat.strain, random=~1|rat.ID, data=my.dataframe)
anova(results.lme)
```

### Two-Way ANOVA with Repeated Measures in dataset2
```{r,cache=TRUE}
set.seed(5250)

myData <- data.frame(PID = rep(seq(from = 1,
                               to = 50, by = 1), 20),
                     stress = sample(x = 1:100,
                                     size = 1000,
                                     replace = TRUE),
                     image = sample(c("Happy", "Angry"),
                                    size = 1000,
                                    replace = TRUE),
                     music = sample(c("Disney", "Horror"),
                                    size = 1000,
                                    replace = TRUE)
)

myData <- within(myData, {
  PID   <- factor(PID)
  image <- factor(image)
  music <- factor(music)
})

myData <- myData[order(myData$PID), ]
head(myData)


#Extracting Condition Means
myData.mean <- aggregate(myData$stress,
                      by = list(myData$PID, myData$music,
                              myData$image),
                      FUN = 'mean')

colnames(myData.mean) <- c("PID","music","image","stress")

myData.mean <- myData.mean[order(myData.mean$PID), ]
head(myData.mean)

#Building the ANOVA
stress.aov1 <- with(myData.mean,
                   aov(stress ~ music * image +
                       Error(PID / (music * image)))
)
summary(stress.aov1)

stress.aov2 <- with(myData,
                   aov(stress ~ music * image +
                       Error(PID / (music * image)))
)
summary(stress.aov2)

stress.aov3 <- with(myData,
                   aov(stress ~ music * image +
                       Error(PID/music))
)
summary(stress.aov3)

stress.aov4 <- aov(stress~music * image + Error(PID/music), data=myData)
summary(stress.aov4)


library(nlme)
stress.lme0 <- lme(stress~music*image, random=~1|PID, data=myData.mean)
anova(stress.lme0)

stress.lme1 <- lme(stress~music*image, random=~1|PID, data=myData)
anova(stress.lme1)

stress.lme2 <- lme(stress~music*image, random=~image|PID, data=myData)
anova(stress.lme2)
```




