---
title: "Linear Mixed-Effects Models by lme4"
author: "test"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



https://cran.r-project.org/web/packages/lme4/index.html


linear mixed-effects models can be determined using the lmer function in the lme4 package for R



# Examples

```{r}
# Example data

library(lme4)
str(sleepstudy)

```


## Simple example: estimates of the standard deviations of the random effects for the intercept and the slope

```{r}

fm1 <- lmer(Reaction ~ Days + (Days | Subject), sleepstudy)

plot(fm1, type = c("p", "smooth"))

```


## resp ~ FEexpr + (REexpr1 | factor1) + (REexpr2 | factor2) + ...


resp ~ FEexpr + (REexpr1 | factor1) + (REexpr2 | factor2) + ...  

(1 | g), is the simplest possible mixed-model formula, where each level of the grouping factor, g, has its own random intercept.


where FEexpr is an expression determining the columns of the fixed-effects model matrix, X,
and the random-effects terms, (REexpr1 | factor1) and (REexpr2 | factor2), determine
both the random-effects model matrix, Z (Section 2.3.2), and the structure of the relative
covariance factor, Λθ


# examples
## Adding group-level predictors in GLMM using lme4

https://biologyforfun.wordpress.com/2017/06/19/adding-group-level-predictors-in-glmm-using-lme4/

```{r}





```



