--- 
title: "Report Example"
author: "Shilin Zhao<br><small>Department of Biostatistics<br>Vanderbilt University School of Medicine</small>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: true
    # toc_float: 
    #   collapsed: true
    code_folding: hide
    # theme: cerulean
    # keep_md: true
description: "some description ..."
---
```{r setup,echo=FALSE}
#require(Hmisc)    # provides knitrSet and other functions
#knitrSet(lang='markdown', fig.path='png/', fig.align='left', w=6.5, h=4.5, cache=TRUE)
# If using blogdown: knitrSet(lang='blogdown')


knitr::opts_chunk$set(echo = TRUE)
options(width = 3000)

#Hmisc package html Special characters issue
options(htmlSpecialType='&')
```


<!-- # Notes -->
<!-- ## Hiding/Moving TOC: -->
<!-- http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate -->

<!-- `r source("https://raw.githubusercontent.com/harrelfe/Hmisc/master/R/hidingTOC.r")` -->
<!-- `r hidingTOC(buttonLabel="Outline")` -->

<!-- ## More themes: -->
<!-- http://www.datadreaming.org/post/r-markdown-theme-gallery/ -->


<!-- # Functions and packages -->

```{r,message=FALSE,warning=FALSE}
library(knitr)
library(ggplot2)
library(RColorBrewer)
library(reshape2)
library(formattable)
library(rms)
library(htmlTable)


#source('D:/source/cqsR/R/summaryTableFunctions.R')
#devtools::source_url("https://raw.github.com/slzhao/cqsR/master/R/summaryTableFunctions.R")
library(cqsR)

```


# Make/Load Data

```{r}
set.seed(123)
rawData<-matrix(nrow=200,ncol=10,rnorm(200*10))
row.names(rawData)<-paste0("Sample",1:200)
colnames(rawData)<-paste0("Feature",1:10)
rawData<-data.frame(rawData,FeatureYN1=sample(c(rep(0,100),rep(1,100)),200),FeatureYN2=as.character(sample(c(rep(0,100),rep(1,100)),200)),FeatureGroup1=as.character(sample(c(rep("Pre/Group1",100),rep("Post/Group2",100)),200)),FeatureCategory1=sample(c("D","E"),200,replace = TRUE),FeatureCategory2=sample(c("A","B","C","D"),200,replace = TRUE),stringsAsFactors = TRUE)

#make some NAs
rawData[c(30:40),"Feature6"]=NA
rawData[c(35:70),"Feature7"]=NA
rawData[c(1:6),"FeatureYN2"]=NA

```

# Table: Paired Test
## Data format: Two columns, Feature6 (Pre) and Feature7 (Post)
```{r,results='asis'}

tableOut=summaryTable(rawData,varCols=c("FeatureYN2","Feature6","FeatureCategory1"),varColsPaired=c("FeatureYN1","Feature7","FeatureCategory2"),pairedTest = TRUE)
printSummaryTable(tableOut)

```

## Data format: One data column (Feature6) and one group column (FeatureGroup1, define Pre and Post)
```{r,results='asis'}

tableOut=summaryTable(rawData,groupCol="FeatureGroup1",varCols=c("FeatureYN2","Feature6","FeatureCategory2"),pairedTest = TRUE)
printSummaryTable(tableOut)

```


# Same result in regular paired tests to make sure the table is correct
```{r}
#First table
matrixForTest=table(rawData$FeatureYN2,rawData$FeatureYN1)
mcnemar.test(matrixForTest)

#Second table
dataOneGroup1=as.character(rawData[which(rawData$FeatureYN1==0),"FeatureYN2"])
dataOneGroup2=as.character(rawData[which(rawData$FeatureYN1==1),"FeatureYN2"])
matrixForTest=table(dataOneGroup1,dataOneGroup2)
mcnemar.test(matrixForTest)

```


# Table: Not Paired Test
## Data format: Two columns, Feature6 (Group 1) and Feature7 (Group 2)
```{r,results='asis'}
tableOut=summaryTable(rawData,varCols=c("FeatureYN2","Feature6","FeatureCategory1"),varColsPaired=c("FeatureYN1","Feature7","FeatureCategory2"))
printSummaryTable(tableOut)

```

## Data format: One data column (Feature6) and one group column (FeatureGroup1, define Group 1 and Group 2)
```{r,results='asis'}

tableOut=summaryTable(rawData,groupCol="FeatureGroup1",varCols=c("FeatureYN2","Feature6","FeatureCategory2"))
printSummaryTable(tableOut)

```

# Table: Same result as Not Paired Test in Hmisc package

```{r,results='asis'}

outVar="FeatureGroup1"
varForTable1=c("FeatureYN2","Feature6","FeatureCategory2")
formulaForTable<-as.formula(paste0(paste(varForTable1, collapse=" + "),"~",outVar))
s <- summaryM(formulaForTable, data=rawData,
							overall=TRUE, test=TRUE)

html(s, exclude1=FALSE,  what=c('%'),digits=3, prmsd=TRUE)

```



# Table: Continuous Group Variable

```{r,results='asis'}
temp<-summaryTableContinus(rawData,variables=c("Feature1","Feature2","FeatureYN1","FeatureCategory2"),groupVariable="Feature3")
printSummaryTableContinus(temp)
```


# Table Example: Showing data.frame in a html table
```{r,results='asis'}
library(htmlTable)
htmlTable(iris[1:3,1:3])


kable(iris[1:3,1:3], 
      caption="A test table")
```


