---
title: "R for Data Science Notes"
author: "Shilin Zhao"
date: "8/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


https://r4ds.had.co.nz/

```{r}
library(tidyverse)
#library(dplyr)
```


# read data
## read excel

```{r}
library(readxl)
readxl_example()

```

## vroom package: Fast reading of delimited files 

https://github.com/r-lib/vroom
https://vroom.r-lib.org/articles/vroom.html



# dplyr package

## join
All columns from x or y will be kept. If multiple matches between x and y, all combinations of the matches are returned:

* inner_join: return all rows from x where there are matching values in y
* left_join: return all rows from x
* right_join: return all rows from y
* full_join: return all rows and all columns from both x and y

Filtering joins keep columns from x:

* semi_join: return all rows from x where there are matching values in y. It will never duplicate rows of x even if one row matching multiple y (as cloumns in y will not be kept, no need to duplicate row in x).
* anti_join: return all rows from x where there are not matching values in y

Nesting joins create a list column of for columns from y:

* nest_join: return all rows and all columns from x. Adds a list column of tibbles. Each tibble contains all the rows from y that match that row of x. 

## filter row based on providing values and use order of values (like selecting by row.names in base R)

```{r}
mtcars1=as_tibble(mtcars,rownames="car")
selectedCars=c("Lotus Europa" ,"Maserati Bora","Ferrari Dino" )

mtcars1 %>% filter(car %in% selectedCars) %>% arrange(match(car,(selectedCars)))
#Or
enframe(selectedCars,name=NULL,value="car")%>% left_join(mtcars1)
data.frame(car=selectedCars,stringsAsFactors = FALSE) %>% left_join(mtcars1)
mtcars1 %>% semi_join(data.frame(car=selectedCars,stringsAsFactors = FALSE))

```

## top_n, top_frac: first rows by a variable value, If n is positive, selects the top rows. If negative, selects the bottom rows. It takes all rows when having tie in top_n. can add %>% slice(1) if only need first one
## slice: first rows by position
## between


# readr package
## read_tsv, read_csv, read_delim


# Functions

## mutate: Make new variable

```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)
```

## mutate+case_when

```{r}
x <- 1:50
case_when(
  x %% 35 == 0 ~ "fizz buzz",
  x %% 5 == 0 ~ "fizz",
  x %% 7 == 0 ~ "buzz",
  TRUE ~ as.character(x)
)
```

## group_by: takes an existing tbl and converts it into a grouped tbl 
```{r}
by_cyl <- mtcars %>% group_by(cyl)

# grouping doesn't change how the data looks (apart from listing
# how it's grouped):
by_cyl

# It changes how it acts with the other dplyr verbs:
by_cyl %>% summarise(
  disp = mean(disp),
  hp = mean(hp)
)


```



## count: Count number per factor

```{r}
#sum (count) of manufacturer
mpg %>% 
  count(manufacturer)

#sum of hwy by manufacturer
mpg %>% 
  count(manufacturer,wt=hwy)

```

## gather:  takes multiple columns and collapses into key-value pairs, or use pivot_longer

```{r}
# From http://stackoverflow.com/questions/1181060
stocks <- tibble(
  time = as.Date('2009-01-01') + 0:9,
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4)
)
head(stocks)

#use everything except "time" as values. "stock" and "price" as new variable names for key and value
gather(stocks, "stock", "price", -time)
#use X, Y, Z as values, so don't need -time. "stock" and "price" as new variable names for key and value
stocks %>% gather("X", "Y","Z",key="stock",value="price")
```

## spread: opposite of gathering. use it when an observation is scattered across multiple rows.

Or use pivot_wider as spread is retired

```{r}
stocksm <- stocks %>% gather(stock, price, -time)
head(stocksm)
stocksm %>% spread(stock, price)
stocksm %>% spread(time, price)


#If duplicate row identifiers and want to keep all of them
#https://github.com/tidyverse/tidyr/issues/426
#https://www.r-bloggers.com/workaround-for-tidyrspread-with-duplicate-row-identifiers/
df=tibble(age=c(21,17,32,38,35),gender=c("Male","Female","Female","Male","Male"))
df <- df %>% 
  group_by(gender) %>% 
  mutate(grouped_id = row_number())
df
df %>% 
  spread(gender, age) %>% 
  select(-grouped_id)
```

## separate: pulls apart one column into multiple columns, by splitting wherever a separator character appears
```{r}
df <- data.frame(x = c(NA, "a.b", "a.d", "b.c"))
df %>% separate(x, c("A", "B"))
# If you just want the second variable:
df %>% separate(x, c(NA, "B"))

#use extra and fill arguments to control what happens
df <- data.frame(x = c("a", "a b", "a b c", NA))
df %>% separate(x, c("a", "b"))
# The same behaviour drops the c but no warnings
df %>% separate(x, c("a", "b"), extra = "drop", fill = "right")

# If only want to split specified number of times use extra = "merge"
df <- data.frame(x = c("x: 123", "y: error: 7"))
df %>% separate(x, c("key", "value"), ": ", extra = "merge")

# convert = TRUE detects column classes
# Argument col can take quasiquotation to work with strings
var <- "x"
df %>% separate(!!var, c("key","value"), ":")
```


## unite: is the inverse of separate(): it combines multiple columns into a single column.
```{r}

unite(mtcars, "vs_am", c("vs","am"))


```

#tidyr package

## separate_rows: Separate a collapsed column into multiple rows

```{r}
library(tidyr)
#If more than one column, should have the same number of rows after seprate
df <- data.frame(
  x = 1:3,
  y = c("a", "d,e,f", "g,h"),
  z = c("1", "2,3,4", "5,6"),
  stringsAsFactors = FALSE
)
separate_rows(df, y, z, convert = TRUE)


```

# purrr package
## reduce: combines the elements of a vector into a single value

# Other notes

## tidyeval 
```{r}
#sym() creates a symbol from a string and syms() creates a list of symbols from a character vector.

df <- data.frame(x = c("a", "a b", "a b c", NA))

var <- "x"
df %>% separate(!!var, c("key","value","note"), " ")

df %>% summarise(n=length(grep(" ",x))) #2
df %>% summarise(n=length(grep(" ",!!var))) #0, as it grep in value of x
df %>% summarise(n=length(grep(" ",!!sym(var)))) #2, correct

```
