---
title: "R Packages Examples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# editData

An RStudio addin for editing a 'data.frame' or a 'tibble'. See <https://github.com/cardiomoon/editData> and <https://cran.r-project.org/web/packages/editData/README.html> .

# rhandsontable

A htmlwidget based on the handsontable.js library. Providing tables in Excel-like appearance and powerful Excel-like features. See <https://jrowen.github.io/rhandsontable/> .


# flexdashboard: easier shiny daashboard, with R markdown integration 

https://rmarkdown.rstudio.com/flexdashboard/



# Clustering for categorical data

https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.476.4707&rep=rep1&type=pdf

CFA, MCA and LCA are well known methods in categorical data analyses. CFA is applied as an instrument in terms of type exploration and confirmation. MCA and LCA are used by means of a more detailed description and 
explanation of statistically signifi cant types.

An example application paper: https://bmcpublichealth.biomedcentral.com/articles/10.1186/s12889-019-6488-8

## CA and MCA analysis in R
http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/
http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/114-mca-multiple-correspondence-analysis-in-r-essentials/



# nomclust: Caculate distance and clustering of categorical data

https://www.researchgate.net/publication/286927797_nomclust_An_R_Package_for_Hierarchical_Clustering_of_Objects_Characterized_by_Nominal_Variables

```{r}

library(nomclust)

data(data20)


```

# klaR: K-Modes clustering (k means to categorical data)

https://dabblingwithdata.wordpress.com/2016/10/10/clustering-categorical-data-with-r/

# cba (ROCK): A robost clustering for categorical data

http://www.inf.ed.ac.uk/teaching/courses/dme/studpres/Rock.pdf
https://dabblingwithdata.wordpress.com/2016/10/10/clustering-categorical-data-with-r/


# poLCA: An R Package for Polytomous Variable Latent Class Analysis: clustering for polytomous outcome variables

https://www.jstatsoft.org/article/view/v042i10

```{r}
library(poLCA)
?poLCA

data(election)
f2a <- cbind(MORALG,CARESG,KNOWG,LEADG,DISHONG,INTELG,
             MORALB,CARESB,KNOWB,LEADB,DISHONB,INTELB)~PARTY
#please note all outcomes are factors

#decide the best number of nclass
resultAll=list()
for (nclass in 3:9) {
  #nes2a <- poLCA(f2a,election,nclass=nclass,nrep=5)
  resultAll[[nclass]] <- poLCA(f2a,election,nclass=nclass,nrep=5)
}
resultAllTable=NULL
for (nclass in 3:9) {
  resultOne=resultAll[[nclass]]
  resultAllTable=rbind(resultAllTable,c(nclass=nclass,DegreeFreedom=resultOne$npar,AIC=resultOne$aic,BIC=resultOne$bic,Gsq=resultOne$Gsq))
}
plot(resultAllTable[,3],type="b")







```


# ggfortify: Data Visualization Tools for Statistical Analysis Results

https://cran.r-project.org/web/packages/ggfortify/index.html

## Plot PCA and factor analysis

https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html

```{r}

library(ggfortify)
df <- iris[1:4]
pca_res <- prcomp(df, scale. = TRUE)

autoplot(pca_res, data = iris, colour = 'Species',loadings = TRUE)

```



# NMF: Run NMF in R

Package: https://cran.r-project.org/web/packages/NMF/vignettes/NMF-vignette.pdf
Method Paper: https://www.pnas.org/content/101/12/4164

##  Use case: Golub dataset
```{r}

library(NMF)

data(esGolub)
esGolub
esGolub <- esGolub[1:200,]

#data format, samples in column, features in row
esGolub@assayData$exprs[1:4,1:4]
dim(esGolub)
#Features  Samples
#     200       38


#Multiple runs
res.multirun <- nmf(esGolub, 3, nrun=5)


```


### Single run
```{r}

res <- nmf(esGolub, 3)
res

#fitted model
fit(res)
#fitted result
V.hat <- fitted(res)
dim(V.hat)
# performance measures and compare with input data
summary(res)
summary(res, target=esGolub)
summary(res, class=esGolub$Cell)

#The basis matrix (i.e. matrix W or the metagenes) and the mixture coefficient matrix (i.e matrix H or the metagene expression profiles)
# get matrix W
w <- basis(res)
dim(w)
# get matrix H
h <- coef(res)
dim(h)

#Extracting metagene-specific features
# only compute the scores
s <- featureScore(res)
summary(s)
# compute the scores and characterize each metagene
s <- extractFeatures(res)
str(s)




```


### Multiple runs
```{r}

res.multirun <- nmf(esGolub, 3, nrun=5)
res.multirun #By default, the returned object only contains the best fit over all the runs
res <- nmf(esGolub, 3, nrun=5, .options=list(keep.all=TRUE)) #can keep all runs


```


### Parallel computations

https://cran.r-project.org/web/packages/NMF/vignettes/NMF-vignette.pdf. Page 17.
Automaticlly. Based on the doParallel and parallel packages. Use .pbackend='mpi' or .pbackend=NULL





# iClusterPlus: Omics Data Integration

https://bioconductor.org/packages/release/bioc/html/iClusterPlus.html

## NMF based
https://compgenomr.github.io/book/matrix-factorization-methods-for-unsupervised-multi-omics-data-integration.html


