---
title: "SPC charts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Thoughts



If I understand it correctly, there are three different (but related) tests in SPC charts:
1.	If a point is a special cause variation: Shewhart’s 3-sigma rule was used in this situation and the point will be labeled by a red dot in qicharts2.
2.	If there is any special cause variation in the whole SPC charts: You mentioned “Western Electric rules” and your “Anhøj rules” in qicharts2 vignette and the center line will be a red dashed line in qicharts2 if this was found by Anhøj rules.
3.	If (and when) we should recompute the limits (Identifying Shifts): and 



# using qicharts2 package and Anhoj rules

https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html

https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-018-0564-0

```{r}
library(tidyverse)
#library(dplyr)
library(qicharts2)


```



# When to make "Shifts" on PSC charts

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6221583/

“eight-point rule” (EPR) (sometimes 7 or 9, depending on the set of rules being used) has become the predominant way to identify a centerline shift—that is, the process has shifted if 8 or more consecutive points occur on the same side of the center line. 


## An test example for qicharts2 package

```{r}

rm(cdi)
library(qicharts2)

i=25:31
cdi$n[i]=cdi$n[i]+8
i=32
cdi$n[i]=cdi$n[i]+5

#There are some special cause variations (red dashed line)
qic(month, n,    notes    = notes,    data     = cdi)

#part=24, the second part still has special cause variations (red dashed line)
qic(month, n,    notes    = notes,    data     = cdi,part=24)

#part=29 gets better result (no red dashed line)
qic(month, n,    notes    = notes,    data     = cdi,part=29)


```

# SPC charts vs regression 

https://www.propharmagroup.com/blog/which-best-regression-versus-control-charting/
