---
title: "limma Test and Design"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://www.bioconductor.org/packages/release/bioc/vignettes/limma/inst/doc/usersguide.pdf


```{r}
library(limma)
```

# MakeDesign and Contrast in One Factor: Two methods

Method 1
```{r}
#####################################
#Two groups
#####################################

design <- cbind(WT=1,MUvsWT=c(0,0,1,1,1))
design
#Or by
Group=factor(c(0,0,1,1,1))
design <- model.matrix(~Group)
colnames(design) <- c("WT","MUvsWT")
design

#fit <- lmFit(eset, design)
#fit <- eBayes(fit)
#topTable(fit, coef="MUvsWT", adjust="BH")


```

Method 2
```{r}
#####################################
#Two groups
#####################################

design <- cbind(WT=c(1,1,0,0,0),MU=c(0,0,1,1,1))
design
#Or By
design <- model.matrix(~0+Group)
colnames(design) <- c("WT","MU")
design

cont.matrix <- makeContrasts(MUvsWT=MU-WT, levels=design)
cont.matrix
#fit <- lmFit(eset, design)
#fit2 <- contrasts.fit(fit, cont.matrix)
#fit2 <- eBayes(fit2)
#topTable(fit2, adjust="BH")

#####################################
#Three groups
#####################################

targets=c("RNA1","RNA1","RNA2","RNA2","RNA3","RNA3")
f <- factor(targets, levels=c("RNA1","RNA2","RNA3"))
design <- model.matrix(~0+f)
colnames(design) <- c("RNA1","RNA2","RNA3")
design

contrast.matrix <- makeContrasts(RNA2-RNA1, RNA3-RNA2, RNA3-RNA1,levels=design)
contrast.matrix

#fit <- lmFit(eset, design)
#fit2 <- contrasts.fit(fit, contrast.matrix)
#fit2 <- eBayes(fit2)
#topTable(fit2, coef=1, adjust="BH")

```


# Examples with Data

```{r}
sd <- 0.3*sqrt(4/rchisq(100,df=4))
y <- matrix(rnorm(100*10,sd=sd),100,10)
rownames(y) <- paste("Gene",1:100)
y[1:2,4:6] <- y[1:2,4:6] + 2
```

## One Factor: Three Groups

```{r}
#Method 1
Treat <- factor(c("C","C","C","T","T","T","G","G","G","G"), levels=c("C","T","G"))
design <- model.matrix(~Treat)
design
fit <- lmFit(y, design)
fit <- eBayes(fit)
TreatT1=topTable(fit, coef="TreatT")
TreatG1=topTable(fit, coef="TreatG")

#Method 2
design <- model.matrix(~0+Treat)
design
fit <- lmFit(y, design)
contrast.matrix <- makeContrasts(TreatTC=TreatT-TreatC,TreatGC=TreatG-TreatC,TreatGT=TreatG-TreatT, levels=design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
TreatT2=topTable(fit2, coef="TreatTC")
TreatG2=topTable(fit2, coef="TreatGC")
TreatGT2=topTable(fit2, coef="TreatGT")

all.equal(TreatT1,TreatT2) #TRUE
all.equal(TreatG1,TreatG2) #TRUE


```


## Two Factors: Paired Data

```{r}
#Method 1
Treat <- factor(c("C","C","C","T","T","T","C","C","T","T"), levels=c("C","T"))
Patient <- factor(c("A","B","C","A","B","C","D","E","D","E"), levels=c("A","B","C","D","E"))
design <- model.matrix(~Treat+Patient)
design
fit <- lmFit(y, design)
fit <- eBayes(fit)
TreatT1=topTable(fit, coef="TreatT")

#Method 2
design <- model.matrix(~0+Treat+Patient)
design
fit <- lmFit(y, design)
contrast.matrix <- makeContrasts(TreatTC=TreatT-TreatC, levels=design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
TreatT2=topTable(fit2, coef="TreatTC")

all.equal(TreatT1,TreatT2) #TRUE


```

## Two Factors: Treatment+Mutation

```{r}
########################################
#Two Factors
########################################
#Method 1
Treat <- factor(c("C","C","C","T","T","T","C","C","T","T"), levels=c("C","T"))
Mutation <- factor(c("Mut","Mut","Mut","Mut","Mut","Mut","WT","WT","WT","WT"), levels=c("WT","Mut"))
design <- model.matrix(~Treat+Mutation)
design
fit <- lmFit(y, design)
fit <- eBayes(fit)
MutationMut1=topTable(fit, coef="MutationMut")

#Method 2
design <- model.matrix(~0+Treat+Mutation)
design
fit <- lmFit(y, design)
contrast.matrix <- makeContrasts(MutationMutWT=MutationMut, levels=design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
MutationMut2=topTable(fit2, coef="MutationMutWT")

all.equal(MutationMut1,MutationMut2) #TRUE





```

## With Interactions: 

Method 2 is good and easy to write interactions

Similar as usersguide 9.5 Interaction Models: 2 × 2 Factorial Designs. But with data to make it easier to reproduce.

### Analysing as for a Single Factor
```{r}
TS <- paste(Mutation, Treat, sep=".")
TS <- factor(TS, levels=c("Mut.C", "Mut.T" ,"WT.C" , "WT.T" ))

design <- model.matrix(~0+TS)
colnames(design) <- levels(TS)
design

fit <- lmFit(y, design)
contrast.matrix <- makeContrasts(
  TvsCinWT=WT.T-WT.C,
  TvsCinMu=Mut.T-Mut.C,
  MuvsWtinC=Mut.C-WT.C,
  MuvsWtinT=Mut.T-WT.T,
  Diff=(Mut.T-Mut.C)-(WT.T-WT.C),
  levels=design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
TvsCinMu1=topTable(fit2, coef="TvsCinMu")
TvsCinWT1=topTable(fit2, coef="TvsCinWT")
MuvsWtinC1=topTable(fit2, coef="MuvsWtinC")
MuvsWtinT1=topTable(fit2, coef="MuvsWtinT")

diff1=topTable(fit2, coef="Diff")


```


### A Nested Interaction Formula

```{r}
#design <- model.matrix(~Mutation+Treat+Mutation:Treat)
design <- model.matrix(~Mutation+Mutation:Treat)
design

fit <- lmFit(y, design)
fit <- eBayes(fit)
TvsCinWT2=topTable(fit, coef="MutationWT:TreatT")
TvsCinMu2=topTable(fit, coef="MutationMut:TreatT")

all.equal(TvsCinWT1,TvsCinWT2) #TRUE
all.equal(TvsCinMu1,TvsCinMu2) #TRUE

# we could extract the interaction contrast Diff considered in "Single Factor" part by
fit2 <- contrasts.fit(fit, c(0,0,-1,1))
fit2 <- eBayes(fit2)
diff2=topTable(fit2)

all.equal(diff1,diff2) #TRUE

```

### Classic Interaction Models

```{r}
design <- model.matrix(~Mutation*Treat)
colnames(design)=gsub(":|\\(|\\)",".",colnames(design))
design

#We need to use contrasts to extract all the comparisons of interest, as we are interested in Mu.S-Mu.U but it is not in design
fit <- lmFit(y, design)
#contrast.matrix <- cbind(TvsCinWT=c(0,0,1,0),TvsCinMu=c(0,0,1,1),Diff=c(0,0,0,1))
contrast.matrix <- makeContrasts(
  MuvsWtinC=MutationMut,
  MuvsWtinT=MutationMut+MutationMut.TreatT,
  TvsCinWT=TreatT,
  TvsCinMu=TreatT+MutationMut.TreatT,
  Diff=MutationMut.TreatT,
  levels=design)

fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)

MuvsWtinC3=topTable(fit2, coef="MuvsWtinC")
MuvsWtinT3=topTable(fit2, coef="MuvsWtinT")
TvsCinWT3=topTable(fit2, coef="TvsCinWT")
TvsCinMu3=topTable(fit2, coef="TvsCinMu")
diff3=topTable(fit2, coef="Diff")

all.equal(TvsCinWT2,TvsCinWT3)
all.equal(TvsCinMu2,TvsCinMu3)
all.equal(diff1,diff3)

all.equal(MuvsWtinC1,MuvsWtinC3)
all.equal(MuvsWtinT1,MuvsWtinT3)



```








